Figures and results for main manuscript

### 2020-11-30 analysis/reports/filtering.Rmd
still figuring out filtering details

@@

ggplot(bbs_statevars, aes(nspecies, nind)) +
  geom_point() +
  geom_hline(yintercept = 40720) +
  geom_vline(xintercept = 200) +
  ggtitle("BBS")

@@

No communities in BBS are outside the maxima. All communities (in this dataset) are only represnted with 1 year.

@@

gentry_statevars <- gentry %>%
  group_by(site) %>%
  summarize(
    nyears = length(unique(year)),
    nspecies = dplyr::n(),
    nind = sum(abund)
  ) %>%
  mutate(outside_maxima = (nspecies > 200) | (nind > 40720))

ggplot(gentry_statevars, aes(nspecies, nind, color = outside_maxima)) +
  geom_point() +
  geom_hline(yintercept = 40720) +
  geom_vline(xintercept = 200) +
  ggtitle("gentry")

@@

Gentry has `r sum(gentry$outside_maxima)` communities, out of `r nrow(gentry)` communities, outside the maxima. RMD just realized, I left these in because they are still sample-able - they have sufficiently low N.

@@

miscabund_statevars <- miscabund %>%
  group_by(site) %>%
  summarize(
    nspecies = dplyr::n(),
    nind = sum(abund),
    nind_na_rm=sum(abund, na.rm = T)
  ) %>%
  mutate(outside_maxima = (nspecies > 200) | (nind > 40720))

ggplot(miscabund_statevars, aes(nspecies, nind, color = outside_maxima)) +
  geom_point() +
  geom_hline(yintercept = 40720) +
  geom_vline(xintercept = 200) +
  ggtitle("miscabund")

@@

There are 130 datasets where there are a mixture of relative and count abundances included. of those, 124 have 0 individuals in the count. THose were not included. There were **6** that have a mixutre and a nonzero number counted. These were INCLUDED in the initial analysis, with relative abundance species completely removed.

miscabund_within <- miscabund_statevars %>%
  filter(nspecies <= 200, nind <= 40720)

miscabund_filtered <- load_dataset("misc_abund_short")

miscabund_sites <- unique(miscabund_within$site)
miscabund_filtered_sites <- unique(miscabund_filtered$site)

included_sites <- as.numeric(miscabund_filtered_sites)

setdiff(included_sites, miscabund_sites)

@@

fia_statevars <- fia %>%
  group_by(site) %>%
  summarize(
    nspecies = dplyr::n(),
    nind = sum(abund),
    nind_na_rm=sum(abund, na.rm = T)
  ) %>%
  mutate(outside_maxima = (nspecies > 200) | (nind > 40720),
         n_equals_s = nind == nspecies)
sum(fia_statevars$n_equals_s)

ggplot(fia_statevars, aes(nspecies, nind, color = outside_maxima)) +
  geom_point() +
  geom_hline(yintercept = 40720) +
  geom_vline(xintercept = 200) +
  ggtitle("fia")

@@

FIA has no extremely large datasets.

@@

ggplot(fia_statevars, aes(nspecies, nind, color = outside_maxima)) +
  geom_point() +
  geom_vline(xintercept = 10) +
  ggtitle("fia")

ggplot(fia_statevars, aes(nspecies)) +
  geom_histogram() +
  geom_vline(xintercept = 10)

@@

mcdb_statevars <- mcdb %>%
  group_by(site) %>%
  summarize(
    nspecies = dplyr::n(),
    nind = sum(abund),
    nind_na_rm=sum(abund, na.rm = T)
  ) %>%
  mutate(outside_maxima = (nspecies > 200) | (nind > 40720),
         n_equals_s = nind == nspecies)

ggplot(mcdb_statevars, aes(nspecies, nind, color = outside_maxima)) +
  geom_point() +
  geom_hline(yintercept = 40720) +
  geom_vline(xintercept = 200) +
  ggtitle("mcdb")

sum(mcdb_statevars$n_equals_s)
```

all_statevars <- miscabund_statevars %>%
  select(-nind) %>%
  rename(nind = nind_na_rm) %>%
  bind_rows(bbs_statevars, fia_statevars, gentry_statevars, mcdb_statevars) %>%
  filter(!outside_maxima) %>%
  mutate(n_equals_s = nind == nspecies)

sum(all_statevars$n_equals_s)

@@

>>>>

The `download_data` function downloads raw data files from https://github.com/weecology/sad-comparison/ (for BBS, Gentry, Mammal Community  Database, and FIA) and figshare http://figshare.com/files/3097079 (for the Miscellaneous Abundance Database). These raw files are stored in `working-data\abund_data` and are not edited.

To filter, we can re-load the raw data files and go through the filtering process step by step. In the analysis this is accomplished by running dataset-specific filtering scripts and saving new .csvs, which are then loaded using `load_dataset`.

We filter according to the following limitations:
* Communities with very large numbers of individuals become computationally intractable. We set the upper limit at 40720, because this is the largest community we were able to sample given the available resources. This upper limit results in the removal of a total of 6 communities across all the datasets.
* Communities for which N = S have only 1 possible SAD.
* Communities with only 1 species have only 1 possible SAD.
* The Miscellaneous Abundance Database includes communities with abundances reported as relative abundances as well as as counts. We exclude communities with relative abundance data. 
* We additionally filter the FIA dataset, because it has many more communities than any other dataset and sampling all of them overwhelms our computational pipeline. We include all communities with more than 10 species (10,000), and an additional, randomly selected, 10,000 (of a remaining 90,000) with 3-10 species. 
* Communities with only 2 species are removed from analyses for skewness, because `e1071::skewness()` always evaluates to 0 for a vector of length 2.

### BBS

@@

ggplot(bbs_statevars, aes(nspecies, nind)) +
  geom_point() +
  ggtitle("BBS")

@@

range(bbs_statevars$nspecies)
range(bbs_statevars$nind)

```
All communities in BBS are used.

### Gentry

@@

gentry_statevars <- gentry %>%
  group_by(site) %>%
  summarize(
    nyears = length(unique(year)),
    nspecies = dplyr::n(),
    nind = sum(abund)
  )

ggplot(gentry_statevars, aes(nspecies, nind)) +
  geom_point() +
  ggtitle("gentry")

range(gentry_statevars$nspecies)
range(gentry_statevars$nind)

sum(gentry_statevars$nspecies == 2)

@@

All Gentry communities are kept. However, `r sum(gentry_statevars$nspecies == 2)` site only has 2 species and will be removed from skewness analyses later in the pipeline.

### Mammal Community Database

mcdb_statevars <- mcdb %>%
  group_by(site) %>%
  summarize(
    nspecies = dplyr::n(),
    nind = sum(abund)
  )

ggplot(mcdb_statevars, aes(nspecies, nind)) +
  geom_point() +
  geom_hline(yintercept = 40720) +
  geom_vline(xintercept = 200) +
  ggtitle("mcdb")

range(mcdb_statevars$nspecies)
range(mcdb_statevars$nind)

`r sum(mcdb_statevars$nspecies == 1)` communities are removed because they have only one species.  Additionally, `r sum(mcdb_statevars$nspecies == 2)
` have 2 species and will be removed from skewness analyses. These removals happen post-pipeline.

### Miscellaneous Abundance Database

@@

miscabund_statevars <- miscabund %>%
  group_by(site) %>%
  summarize(
    nspecies = dplyr::n(),
    nind = sum(abund),
    nind_na_rm=sum(abund, na.rm = T)
  ) %>%
  mutate(outside_maxima = (nind_na_rm > 40720))

ggplot(miscabund_statevars, aes(nspecies, nind_na_rm, color = outside_maxima)) +
  geom_point() +
  geom_hline(yintercept = 40720) +
  geom_vline(xintercept = 200) +
  ggtitle("miscabund")

@@

There are 130 datasets where there are a mixture of relative and count abundances included. of those, 124 have 0 individuals in the count. THose were not included. There were **6** that have a mixutre and a nonzero number counted. These were INCLUDED in the initial analysis, with relative abundance species completely removed. Upon realizing this we remove these 6 communities from the final analysis. Rather than re-run the whole pipeline, we manually remove them when we load the data to make the histograms and summary tables (see manuscript_main.Rmd).

### FIA

@@

fia_statevars <- fia %>%
  group_by(site) %>%
  summarize(
    nspecies = dplyr::n(),
    nind = sum(abund),
    nind_na_rm=sum(abund, na.rm = T)
  )

ggplot(fia_statevars, aes(nspecies, nind)) +
  geom_point() +
  geom_hline(yintercept = 40720) +
  geom_vline(xintercept = 200) +
  ggtitle("fia")

@@

FIA has no extremely large datasets. However, it has `r nrow(fia_statevars)` communities, of which `r sum(fia_statevars$nspecies < 10)` have fewer than 10 species. This many communities overwhelms our computational pipeline. We therefore sample all `r sum(fia_statevars$nspecies >= 10)` communities with 10 or more species, and a random subsample of 10,000 communities with 3-9 species. We then run these through the pipeline as two separate databases. `fia_short` is the communities with 10 or more species, and `fia_small` is the 10,000 communities with 3-9 species.

To go through this we re-load FIA using load_dataset, so we can get the `site` column in the correct format and order such that it will replicate when we run sample.

@@

fia <- load_dataset("fia")

fia_sv_short <- get_statevars(fia) %>%
    dplyr::filter(s0 >= 10)

fia_short <- fia %>%
  dplyr::filter(site %in% fia_sv_short$site) %>%
  dplyr::mutate(dat = "fia_short")

fia_short_statevars <- fia_short %>%
  group_by(site) %>%
  summarize(
    nspecies = dplyr::n(),
    nind = sum(abund),
    nind_na_rm=sum(abund, na.rm = T)
  ) %>%
  mutate(site = as.numeric(site))

ggplot(fia_short_statevars, aes(nspecies, nind)) +
  geom_point() +
  ggtitle("fia short")

fia_sv_small <- get_statevars(fia) %>%
  dplyr::filter(s0 >= 3) %>%
  dplyr::filter(s0 <= 9)

  set.seed(1977)
  fia_sv_small <- fia_sv_small[ sample.int(nrow(fia_sv_small), size = 10000, replace = F), ]

fia_small <- fia %>%
  dplyr::filter(site %in% fia_sv_small$site) %>%
  dplyr::mutate(dat = "fia_small")

fia_small_statevars <-fia_small  %>%
  group_by(site) %>%
  summarize(
    nspecies = dplyr::n(),
    nind = sum(abund),
    nind_na_rm=sum(abund, na.rm = T)
  )  %>%
  mutate(site = as.numeric(site))

ggplot(fia_small_statevars, aes(nspecies, nind)) +
  geom_point() +
  ggtitle("fia small")

fia_small_loaded <- load_dataset("fia_small")

setdiff(fia_small_loaded$site, fia_small$site)

@@

all_statevars <- miscabund_statevars %>%
  select(-nind) %>%
  rename(nind = nind_na_rm) %>%
  bind_rows(bbs_statevars, fia_short_statevars, fia_small_statevars, gentry_statevars, mcdb_statevars) %>%
  mutate(one_species = nspecies == 1,
         two_species = nspecies == 2,
         n_equals_s = nspecies == nind)

@@

Overall there are `r sum(all_statevars$one_species)` removed at the end because they have just 1 species. There are `r sum(all_statevars$two_species)` that will be removed from skewness because they ahve 2 species. There are `r sum(all_statevars$n_equals_s)` that are removed because N = S.

Note that, in practice, many more small ones end up getting removed because they have very few possible SADs and very few unique values for skewness and evenness.

```{r}
all_di <- read.csv(here::here("analysis", "reports", "all_di.csv"), stringsAsFactors = F)

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 > s0,
         s0 > 1,
         !singletons,
         dat %in% c("bbs", "fia_short", "fia_small", "gentry", "mcdb", "misc_abund_short")) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
    Dataset = ifelse(Dataset == "fia", "Forest Inventory and Analysis", Dataset),
        Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
        Dataset = ifelse(Dataset == "mcdb", "Mammal Community DB", Dataset),
        Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
        Dataset = ifelse(Dataset == "misc_abund", "Miscellaneous Abundance DB", Dataset))

nrow(all_di)

filtered_sv <- all_statevars %>%
  filter(!n_equals_s,
         !one_species)

nrow(filtered_sv)
```

### 2020-11-30 analysis/reports/manuscript_main.Rmd
continuing to t r y
still figuring out filtering details

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 > s0,
         s0 > 1,
         !singletons,
         dat %in% c("bbs", "fia_short", "fia_small", "gentry", "mcdb", "misc_abund_short")) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
    Dataset = ifelse(Dataset == "fia", "Forest Inventory and Analysis", Dataset),
        Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
        Dataset = ifelse(Dataset == "mcdb", "Mammal Community DB", Dataset),
        Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
        Dataset = ifelse(Dataset == "misc_abund", "Miscellaneous Abundance DB", Dataset))
```

>>>>

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 > s0,
         s0 > 1,
         !singletons,
         dat %in% c("bbs", "fia_short", "fia_small", "gentry", "mcdb", "misc_abund_short")) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
    Dataset = ifelse(Dataset == "fia", "Forest Inventory and Analysis", Dataset),
        Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
        Dataset = ifelse(Dataset == "mcdb", "Mammal Community DB", Dataset),
        Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
        Dataset = ifelse(Dataset == "misc_abund", "Miscellaneous Abundance DB", Dataset))

miscabund_problems <- c(144, 145, 145, 146, 147, 178, 446)

all_di <- all_di %>%
  mutate(is.miscabund.problem = ifelse(dat == "misc_abund",
                                       ifelse(site %in% miscabund_problems, T, F), F)) %>%
  filter(!is.miscabund.problem) %>%
  ungroup() %>%
  select(-is.miscabund.problem)
